---
title: "Project #3: Escapades in Market Risk"
output: flexdashboard::flex_dashboard
---

```{r include=FALSE}
# ----------------------- Background section -------------------------
```

Background {data-orientation=columns}
=====================================  
    
Column {data-width=600}
-------------------------------------
    
### Problem

[Problem statement]

### Additional details

[project resources and info]

### Key business questions

[business questions]

Column {data-width=400}
-------------------------------------
   
### Getting a response: More detailed questions

[Detailed questions]

```{r include=FALSE}
# --------------------- Data Exploration section -----------------------
```

Data Exploration {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------

### Graphs

Metals Market Percent Changes
```{r}
# Add graphs
```
<br/><br/>

Metals Market Percent Changes - Size  
```{r}
# Add graph of size
```
   
### Data

Descriptive Statistics
```{r}
# Add min/1st qu/median/srd qu/max of nickel/copper/alum
```
<br/><br/>

Data Moments
```{r}
# Output the copper.size, alum.size, nickel.dir and copper.dir
# this should be mean, median, std_dev, IQR, skewness, kurtosis
```

Row {data-height=400}
-------------------------------------
    
### Data Insights

[Add data insights from graphs and data]
<br/><br/>

[Talk abut time frame, volatility, correlation, skewness andkurtosis]


```{r include=FALSE}
# --------------------- Volatility section -----------------------
```

Volatility {data-orientation=rows}
=====================================     
   
Row {data-height=300}
-------------------------------------

### Returns % change - ACF

Returns % change - ACF
```{r}
# Add graphs for Returns % change - ACF
```

### Size Returns % change - ACF

Size Returns % change - ACF
```{r}
# Add graphs for Size Returns % change - ACF
```

Row {data-height=300}
-------------------------------------
    
### Returns % change - PACF

Returns % change - ACF
```{r}
# Add graphs for Returns % change - PACF
```

### Size Returns % change - PACF

Size Returns % change - ACF
```{r}
# Add graphs for Size Returns % change - PACF
```

Row {data-height=300}
-------------------------------------

### Data Insights

[Add data insights. Something about weekly vs monthly, tails, volatility, etc ]


```{r include=FALSE}
# --------------------- Correlation Sensitivity section -----------------------
```

Correlation Sensitivity {data-orientation=rows}
=====================================     
   
Row {data-height=700}
-------------------------------------

### nickel-copper correlation sensitivity to copper volatility

nickel-copper correlation sensitivity to copper volatility
```{r}
# Add graphs for nickel-copper correlation sensitivity to copper volatility
```

### copper-aluminum correlation sensitivity to aluminum volatility

copper-aluminum correlation sensitivity to aluminum volatility
```{r}
# Add graphs for copper-aluminum correlation sensitivity to aluminum volatility
```

### nickel-aluminum correlation sensitivity to aluminum volatility

nickel-aluminum correlation sensitivity to aluminum volatility
```{r}
# Add graphs for nickel-aluminum correlation sensitivity to aluminum volatility
```

Row {data-height=300}
-------------------------------------

### Data Insights

[Add data insights. Talk about variance, spikes, positive vs negative]

```{r include=FALSE}
# --------------------- Expected Shortfall section -----------------------
```

Expected Shortfall {data-orientation=rows}
=====================================     
   
Row {data-height=700}
-------------------------------------

### Nickel Shortfall

Nickel Shortfall
```{r}
# Add graphs for Nickel Shortfall
```

### Copper Shortfall

Copper Shortfall
```{r}
# Add graphs for Copper Shortfall
```

### Aluminum Shortfall

Aluminum Shortfall
```{r}
# Add graphs for Aluminum Shortfall
```

Row {data-height=300}
-------------------------------------

### Data Insights

[Add data insights. Which has highest risk, tails, expected shortfall values and value at risk values, what do they mean?]

```{r include=FALSE}
# --------------------- Losses section -----------------------
```

Losses {data-orientation=rows}
=====================================     
   
Row {data-height=700}
-------------------------------------

### Losses Graph

Losses Graph
```{r}
# Add graphs for Losses
```

Row {data-height=300}
-------------------------------------

### Data Insights

[Add data insights. What is the loss limit, VAR, ES, what do they mean?  What about the distribution?]
